A Motion Blur Resilient Fiducial Marker For Quadcopter Imaging

### 2014-06-09 WACV/paper.tex
Actually deleted files from Git

\begin{abstract}
This paper describes the design and evaluation of a binary fiducial marker
for use with low-cost quadcopters.  Fiducial markers are commonly placed
in environments to provide a uniquely identifiable object in the scene.
For quadcopter applications, fiducials are often used for evaluating planning
algorithms by allowing ground truth positions in the scene to be detected from
the quadcopter's camera. Quadcopters, however, are subject to quick and
unstable motions that can cause significant motion blur that severely affects
the detection rate of existing fiducial markers. This problem motivated us to
design a fiducial that is robust to motion blur. Our proposed fiducial design
uses concentric circles with the observation that the direction perpendicular
to the motion blur direction will be relatively unaffected by the blur. This
allows the fiducial code along the blur direction to be recognized.
The circular design allows this to work for all motion blur directions. We
detail the design and detection algorithm for our fiducial and show that our
marker can significantly outperform existing fiducial markers in scenes
captured with a quadcopter.
\end{abstract}

@@

The recent availability of low-cost quadcopters have helped to fuel significant
efforts in research focused on unmanned aerial vehicles. Navigation and planning of
these vehicles is typically performed using onboard inertial sensors and/or
vision based modules that uses visual cues in the real world~(e.g.,
\cite{Davison:2007,Engel12,Engel13}). However, to evaluate the effectiveness of
navigation methods, fiducial markers are commonly placed into the environment
to provide additional information that serves for ground-truth positional
measurements (e.g., \cite{Bosnak:2012,Lim09,Klopschitz:2007}).

In order to be effective, fiducial markers (or simply, fiducials) need
to be easily detected in the scene. A variety of fiducial markers
have been proposed in the literature
(e.g.,~\cite{NaimarkF02,ARToolkit02,Fiala05,Pitag13,runetag11}).
These take the form of binary codes arranged into rectangular grids (\cite{ARToolkit02,Fiala05})
or other geometric primitives arranged in predefined spatial patterns
(\cite{NaimarkF02,Pitag13,runetag11}).
Figure \ref{fig:teaser}-(a) shows the popular ARTag \cite{Fiala05} as
seen from a quadcopter. A problem for existing fiducials is that low-cost quadcopters
often exhibit very quick and erratic physical movements that
result in motion blur in the quadcopter's onboard camera. This motion blur has
an adverse effect on the recognition of fiducial markers. This can be seen in
Figure \ref{fig:teaser}-(b) where the ARTag cannot be recognized due to motion
blur. This is not too surprising as most existing fiducials are not
designed to handle motion blur.

Compounding this problem is the additional issue of dropped video frames from
the quadcopter's wireless communication module. This means that not only is
blur a problem, but there may be large discontinuities in the pattern's
position due to missing video frames. As we will show in Section 4, this later problem
makes it challenging to apply tracking algorithms that can exploit temporal
coherence for determining the fiducial's position.

\noindent\textbf{Contribution}~~To address these problems, we propose a
fiducial that is designed to be resistant to motion blur. Our design is based on
concentric circles as shown in Figure \ref{fig:teaser}-(c,d). The design is based
on the observation that motion blur from a quadcopter tends to be linear in
nature. As such, when our fiducial is blurred, there is no blur in the direction
perpendicular to the direction of motion.  This allows the signature of the
fiducial to remain intact in any direction. In addition, by using concentric
rings, we can treat the presence or absence of a ring as a bit, allowing us to
assign a code to the marker. Our experiments show that this design can
significantly outperform existing codes in the presence of motion blur. As far
as we are aware, this is the first work to propose a blur resistant marker.

The remainder of this paper is organized as follows.  Section 2 gives an
overview of related work related to fiducials as well as the related problem of
tracking. Section 3 discusses our design and detection algorithm as well as
the performance of existing codes under motion blur. Section 4 shows several
experiments using quadcopter imagery. This is followed by a discussion and
summary in Section 5 and 6 respectively.

\begin{figure}
\includegraphics[width=\linewidth]{teaser.pdf}
\caption{This figure shows the experimental setup and comparison of
the output of ALVAR~\cite{alvar} (for detecting ARTag) versus our blur resistant
fiducial in a normal and blurred video frame. The scene contains two ARTags and our code.
(a) shows that the ARTag can be detected when the image has no blur.
(b) shows that ARTags are no longer detectable due to the blur.
(c-d) show the same images in (a) and (b) where our code is being detected.}
\label{fig:teaser}
\end{figure}

@@

\begin{figure}
\centering
 \begin{subfigure}[b]{0.23\linewidth}
  \centering
  \includegraphics[width=0.8\linewidth]{intersense.jpg}
  Circular Data Matrix~\cite{NaimarkF02}
 \end{subfigure}\quad
 \begin{subfigure}[b]{0.14\linewidth}
 \centering
  \includegraphics[width=\linewidth]{pattKanji.pdf}
  ARToolkit~\cite{ARToolkit02}
 \end{subfigure}\quad
 \begin{subfigure}[b]{0.12\linewidth}
  \centering
  \includegraphics[width=\linewidth]{ARtag.jpg}
  ARTag\quad~\cite{Fiala05}
 \end{subfigure}\quad
 \begin{subfigure}[b]{0.12\linewidth}
  \centering
  \includegraphics[width=\linewidth]{pifiducial.jpg}
  PiTag\quad~\cite{Pitag13}
 \end{subfigure}\quad
 \begin{subfigure}[b]{0.14\linewidth}
  \centering
  \includegraphics[width=\linewidth]{our_fiducial}
  Our Fiducial
 \end{subfigure}
 \caption{This figure shows the design of several existing fiducial markers.  While
 each code has their own pros and cons for various environments, none are
 designed to be recognized under motion blur.}
 \label{fig:previous_work}
\end{figure}

\noindent{\textbf{Fiducials}}~Figure~\ref{fig:previous_work} shows several
examples of existing fiducial markers.  Many designs use a two dimensional
barcode inside a rectangular grid. One example of such a fiducial is from the
ARToolkit~\cite{ARToolkit02}, a well known toolkit used in many augmented
reality (AR) applications. Kato and Billinghurst~\cite{kato-artoolkit}
first demonstrated the use of ARToolkit in various augmented-reality-based applications.

Fiala~\cite{Fiala05} proposed a fiducial termed, ARTag, which is a
bi-tonal system consisting of a square border and an interior 6$\times$6 grid
of black or white cells. The improvement of the ARTag compared to the ARToolkit was
detection of corners instead of detection of lines to find possible pattern.
This proved to be more efficient than \cite{ARToolkit02} in terms of
recognition rate as well as the number of different patterns which can be
created.   The reliance on both line and corner detection, however, hampers recognition
under motion blur.

There were also attempts to use circular patterns instead of rectangular.
Gatrell et al.~\cite{concentric} used concentric circles
for monocular pose estimation as well as object identification. Cho et al.
\cite{Cho:2001,Cho97fastcolor} have used multicolor rings instead of black and white rings~\cite{concentric} to increase possible number of fiducials.
These multicolor rings are used in wide area tracking in large scale
applications.  While based on a concentric rings, these approaches require
the full ring to be recognized which is not possible when the pattern undergoes directional
motion blur.

Naimark and Foxlin~\cite{NaimarkF02} proposed a circular bar code
called the Circular Data Matrix that is beneficial in terms of
easy detection and ability to have a large number
of uniquely identifiable codes.  To address the issue of occlusion,
Bergamasco et al.~\cite{runetag11} proposed the RUNE-tag fiducial by dividing
into number of circular dots arranged in circular fashion. RUNE-tags can be
detected even when up to 50\% of fiducial area is occluded. Bergamasco et
al.~\cite{Pitag13} proposed the PiTag fiducial, also composed of circular
structures but arranged in a rectangle to exploit projective invariant cross-ratio.
This provided similar occlusion resistance as RUNE-tag but with even less
circular dots. All of these techniques however rely on generic features
detection (e.g., circle detection) that break down under motion blur.

@@

\noindent{\textbf{Tracking}}~~Fiducial detection between successive video
frames can be considered as a tracking problem where the tracked object is
the fiducial.  There is a very large body of research dedicated to tracking and interested readers are referred to~\cite{Yilmaz:2006} for a good survey.

Most tracking methods (e.g.,~\cite{Ross:2008,Wu:2009,Perez02,Mei:2009}) assume
the image sequence to be blur free. In reality, however, the presence of motion blur in
a video sequence is often unavoidable. To this end, Wu et al.~\cite{Wu:2011}
proposed the Blur-driven Tracker (BLUT) framework for tracking motion-blurred targets. BLUT
is based on the observation that although motion blur degrades the visual
features of the target, it may also provide useful cues about the
movements to help tracking.

The BLUT framework successfully tracks a blurred target when there is uniform motion
and the position of the tracked object does not change drastically in successive
frames. However, as we will demonstrate in Section 4, the erratic motion from the quadcopters as well as the problem of dropped video frames is too problematic to be successively tracked.

@@

We begin by first motivating the need for a new blur resistant fiducial by examining
the performance of prior fiducials under motion blur.  After this, we detail
our design as well as the detection algorithm used to find our marker in an image.

\subsection{Examining Prior Fiducials Under Motion Blur}\label{sec:blurtest}

Here we examine the the performance of two
popular fiducials under motion blur.  Specifically we examine ARTags
~\cite{Fiala05} and PiTags~\cite{Pitag13} given their difference in geometric
design and the availability of API to develop applications to recognize the tags.
RUNE-tag~\cite{runetag11} currently does not provide access to the an
implementation to generate or recognize its tags.  The circular data
matrix~\cite{NaimarkF02} is available as a commercial product, however it
requires proprietary hardware to recognize the fiducial.

To evaluate the ARTag and PiTag under blur, we simulate the appearance of the
marker as seen by a quadcopter by scaling the tags to 150$\times$150 pixels.
Both fiducials are then blurred using linear motion blur at various
orientations with different blur scales. The blur motion ranged from 15 to 50
in magnitude (denoted in pixels), representing small to significant motion
blur. Figure \ref{fig:artag_pitag} shows the visual appearance of the blurred
tags. We then try to detect the markers using the ALVAR library~\cite{alvar} and
PiTag library~\cite{ros_pitag}. The table in the right side of
Figure~\ref{fig:artag_pitag} shows the recognition rate (in percentage) of two
fiducial markers at various blur scales over all the different orientations.
As we can see, the PiTag performance quickly diminishes under small amounts of
blur, while at 35 motion blur, the ARTag's recognition rate drops to less than
20\%.

\noindent\begin{figure}
\includegraphics[width=\linewidth]{artag_pitag.pdf}
%\captionof{figure}{Blurred ARTag and Pi--tag with various blur scales}
%\captionof{table}{Recognition rate of ARTag and Pi-tag fiducial marker atvarious blur scales}
\begin{tabularx}{\linewidth}{|Y|Y|Y|}
\cline{1-3}
\small{Blur} & \multicolumn{2}{c|}{ \small{Recognition Rate}}
\\\cline{2-3}
\small{Scale}& \small{PiTag} &	\small{ARTag} \\ \cline{1-3}
\small{15} & \small{100} & \small{100} \\ %\cline{1-3}
\small{30} & \small{0} & \small{100} \\  %\cline{1-3}
\small{35} & \small{0} & \small{19} \\ %\cline{1-3}
\small{50} & \small{0} & \small{0} \\ \cline{1-3}
\end{tabularx}
\captionof{figure}{Figure showing PiTag and ARTag blurred with various blur
scales at different orientations. Table on the right shows the recognition rate
(in percent) of both fiducial markers at various blur scales along all blur
orientations.  Recognition rates for both tags are significantly reduced due to
blur. For severe blur, detection is not possible.}
\label{fig:artag_pitag}
\end{figure}

@@

We have designed a binary coded fiducial that uses concentric white rings of
equal widths on a black background with a blurred border\footnote{This design
can be inverted to have a white background with white rings.}. The
outermost and innermost rings represent the start and end of the code and is
embedded in the fiducial.  The binary code is represented by the presence (or
absence) of rings between ``marker'' rings.

\begin{figure}
\centering
  \includegraphics[width=.24\linewidth]{newconcentric_00.pdf}
  \includegraphics[width=.24\linewidth]{newconcentric_01.pdf}
  \includegraphics[width=.24\linewidth]{newconcentric_10.pdf}
  \includegraphics[width=.24\linewidth]{newconcentric_11.pdf}
  \caption{Two bit binary coded fiducials (from left to right: binary code 00,
  binary code 01, binary code 10, binary code 11.)}
  \label{fig:fiducials}
\end{figure}

Depending on which ring is present or absent, the resulting binary code will
change. The number of different patterns depends on the number of bits in the
binary code. For example, if the binary code has three bits, there will be a
maximum of three rings between marker-rings and we end up with eight
different patterns.  Figure \ref{fig:fiducials} shows two bit binary coded
fiducials.

\subsection{Fiducial Detection Algorithm}

\begin{figure}
\centering
\includegraphics[width=\linewidth]{blur_direction.pdf}
\caption{Figure showing how change in blur direction changes the location of
unblurred linear pattern.}
\label{fig:blur_direction}
\end{figure}

Our fiducial detection strategy is different from \cite{NaimarkF02,Pitag13} and works
under significant amounts of blur.   As previously mentioned, our approach works
under the observation that the motion blur for the quadcopter's camera
can be well modelled as linear motion.  This  linear motion blur assumption has
been shown to be reasonable in prior works targeting camera motion blur
(e.g.~\cite{Moshe:2003,Moshe:2004}). Under this assumption, the scene content
perpendicular to the blur direction is unaffected by the blur.  Because of our
circular design, the direction perpendicular to the linear motion will still be
recognizable as a linear pattern.  Figure \ref{fig:blur_direction} shows
examples of this using a pattern under various motion directions.

Figure \ref{fig:overall_flow} shows the process involved in fiducial
detection. We give a brief overview of our algorithm here with each step described
in detail afterwards.  Our detection algorithm has four steps. Step 1,
we apply a Gabor filter on the image to isolate the
potential locations of the pattern.  Step 2, we find the clusters of patches
found in the Gabor output by clustering.  Step 3, performs Principal
Component Analysis (PCA) on each cluster to find the dominant direction
unaffected by blur.  Step 4, based on the direction we extract the intensity
profile of the pattern and classify the code using a K-Nearest Neighbor
classifier based on training data.

\begin{figure*}[ht!]
\centering
\includegraphics[width=0.8\linewidth]{overall_flow.pdf}
\caption{An overview of our algorithm to detect and classify fiducial.
The four step process includes (Step 1) Gabor filtering,
(Step 2) component clustering, (Step 3) PCA to determine the dominant direction,
and (Step 4) K-NN classification using training-data captured for each pattern.}
\label{fig:overall_flow}
\end{figure*}

\noindent Details of each of the steps are as follows:\\
\noindent\textbf{Gabor filter}~~A 2D Gabor filter is a Gaussian kernel function
modulated by a sinusoidal plane wave~\cite{Kruizinga:2002}. It is used to find
high gradient patches from the image. In our case, it is used to detect portions
of the circular fiducials that were not affected by the blur.
We applied the Gabor filter in eight
different orientations ($\theta = 0, 45, 90, \ldots ,225, 270, 315$).  The
following parameters were used for creating each Gabor kernel: $\lambda$ (wavelength) $= 8$, $\gamma$
(aspect ratio) $= 0.5$, $\sigma$ (spread) $= 0.56\lambda$, $\psi$
(phase angle) $= 0$ (for real part), $\pi/2$ (for imaginary part).
Then $\ell^2$-norm of outputs along all orientations is calculated and finally
$\ell^2$-normalized image is thresholded with threshold set to $0.4$ (on scale of zero to
one).  For further details to the Gabor filter, see~\cite{Kruizinga:2002}.
%% MBS - is the Kruizinga filter resonable to cite?

\noindent\textbf{Clustering}~~The binarized Gabor filter responses are
treated as a set of connected components in the image.   Clustering
is used to find components that are located in a close spatial region.  We do
this via hierarchical clustering~\cite{ALGLIB} using unweighted
average linkage with a distance threshold set to 150.

\noindent\textbf{PCA}~~For each clustered set of components, we apply
PCA to determine the dominant direction.  This is done by examining the
orientation of the first principal component.  We then extract an intensity
profile patch in the input image along this direction as it extends through the
bounding box of the cluster. The signature of this profile will be used to
identify the code.

After finding the intensity profile, we examine it by
projecting the pixel intensities onto the PCA direction and record the
number and width of the white-to-black transitions.  If the number of transitions or the
width of the transitions are not consistent with what is allowable by
our code, we reject the clustered region as a potential fiducial.
Clustered components that have allowable transitions counts and transitions with
uniform widths are further considered for classification to determine
the binary code.

\noindent\textbf{Classification}~~As mentioned above, a small image patch
containing the intensity profile of the fiducial is used to identify the code.   We found
that a training-based method using K-NN gave better results than trying to
find binary code directly by determining the presence (or absence) of ring
at particular positions in the image patch. To build the training-data, a
synthetic fiducial pattern is blurred along 36 orientations (0, 10, 20, \ldots
, 350) with blur scale set to 40, and its intensity profile along the first
principal component from every output is taken as training data for that
fiducial pattern. Figure \ref{fig:training_data} shows the process of creating
training data for fiducial with binary code ``01'' embedded in it.  For an
input image patch, we normalize its intensity range and then compare it against
the training-data. The class label from the closest top $K=5$ images in the
training-data is used to label the patch.

\begin{figure}[h!]
\centering
  \includegraphics[width=0.95\linewidth]{training_data.pdf}
  \caption{Process to create training data for the fiducial pattern with binary
  code ``01''. The synthetic pattern is blurred along various orientations. Gabor
  filter is applied on the blurred pattern. Intensity profile is found along the
  first principal component of the clustered Gabor output. Same process is used
  to create training data for other fiducial patterns.}
  \label{fig:training_data}
\end{figure}

@@

We have implemented our algorithm in C++ using the OpenCV library.
Experiments were performed on a PC with Intel Core i7 processor(@3.4GHz) and 4GB RAM.
Source code to generate and recognize the code, as well as the data sets used in
this paper will be made publicly available.
%Video clips of collected data are provided in the supplementary material.

>>>>

\begin{abstract}

  Fiducial markers are commonly placed in environments to provide a
  uniquely identifiable object in the scene.  In quadcopter
  applications, these fiducials are often used to evaluate planning
  algorithms given that ground truth positions can be
  detected from the quadcopter's camera.

  Low cost quadcopters, however, are subject to quick and unstable
  motions that can cause significant motion blur that severely affects
  the detection rate of existing fiducial markers. This problem
  motivated us to design a fiducial that is robust to motion blur. Our
  proposed design uses concentric circles with the observation that
  the direction perpendicular to the motion blur direction will be
  relatively unaffected by the blur. As a result, an appropriate
  fiducial code orthogonal to the blur direction can be
  recognized. Since the direction of motion blur is unknown, the
  circular design is good for all motion blur directions. We describe
  the design of binary fiducial markers, and also a detection
  algorithm.  We show that our marker can significantly outperform
  existing fiducial markers in scenes captured with a quadcopter.

\end{abstract}

@@

The recent availability of low-cost quadcopters has helped  fuel
significant efforts in research focused on unmanned aerial
vehicles. Navigation and planning
of these vehicles is typically~\cite{Davison:2007,Engel12,Engel13}
performed using onboard inertial sensors and/or vision based modules
that uses visual cues in the real world. However, to evaluate the
effectiveness of navigation methods, one or more fiducial markers are commonly
placed~\cite{Bosnak:2012,Lim09,Klopschitz:2007}
 in the environment to provide additional information that
serves for ground-truth positional measurements.

In order to be effective, fiducial markers (or simply, fiducials) need
to be easily detected in the scene. A variety of fiducial markers have
been proposed~\cite{NaimarkF02,ARToolkit02,Fiala05,Pitag13,runetag11}
in the literature.  These take the form of binary codes arranged into
rectangular grids~\cite{ARToolkit02,Fiala05} or other geometric
primitives arranged in predefined spatial
patterns~\cite{NaimarkF02,Pitag13,runetag11}.  Figure \ref{fig:teaser}
shows the popular ARTag \cite{Fiala05} fiducial as possibly seen from
a quadcopter. A problem for existing fiducials is that low-cost
quadcopters often exhibit very quick and erratic physical movements
that result in motion blur in the quadcopter's onboard camera. This
motion blur has an adverse effect on the recognition of fiducial
markers. This can be seen in Figure \ref{fig:teaser}--(b) where the
ARTag fiducial cannot be recognized due to motion blur. This is not
too surprising as most existing fiducials are not designed to handle
motion blur.

Compounding this problem is the additional issue of dropped video
frames from the quadcopter's wireless communication module. This means
that not only is blur a problem, but there may be large
discontinuities in the pattern's position due to missing video
frames. As we will show in Section~4, this later problem makes it
challenging to apply tracking algorithms that can exploit temporal
coherence for determining the fiducial's position.

\noindent\textbf{Contributions:}~~To address these problems, we propose a
fiducial that is designed to be resistant to motion blur. Our design
is based on circles as shown in Figure~\ref{fig:teaser}--(c,d). The
design is based on the observation that motion blur from a quadcopter
tends to be linear in nature. As such, when our fiducial is blurred,
there is no blur in the direction perpendicular to the direction of
motion.  This allows the signature of the fiducial to remain intact in
any direction.

When multiple fiducials are present, using concentric rings, we can
treat the presence or absence of a ring as a bit, allowing us to
assign a code to the marker. Our experiments show that these designs can
significantly outperform existing codes in the presence of motion
blur. As far as we are aware, this is the first work to propose a blur
resistant marker especially in quadcopter settings.

The remainder of this paper is organized as follows.  Section~2 gives
an overview of related work in fiducials, as well as the related
problem of tracking. Section~3 motivates our fiducial design by
analysing the performance of existing codes under motion blur, and
describes our detection algorithm. Section~4 shows several
experiments using quadcopter imagery. This is followed by a discussion
and a summary in Section~5 and Section~6 respectively.

\begin{figure}[t!]
  \includegraphics[width=\linewidth]{images/teaser}
  \caption{Experimental setup and comparison of fiducial-based
    algorithms. A green border signifies success. 
    \textbf{(Left)} A drone encounters three different fiducials. 
    \textbf{(Right, Top)} Output of ALVAR~\cite{alvar} under favorable and blurred
    circumstances. No green border is seen in (b) signifying failure.
    \textbf{(Right, Bottom)} Output of the fiducial proposed in this paper. Green
    border is seen in both (c) and (d).
    \label{fig:teaser}}
  \end{figure}

@@

\textbf{Fiducials:}~Figure~\ref{fig:previous_work} shows a few
examples of existing fiducial markers.  Many designs use a two
dimensional barcode inside a rectangular grid. One example of such a
fiducial is from the ARToolkit~\cite{ARToolkit02}, a well known
toolkit used in many augmented reality (AR) applications. Kato and
Billinghurst~\cite{kato-artoolkit} first demonstrated the use of
ARToolkit in various augmented-reality-based applications.

\begin{figure}[b!]
\centering
 \begin{subfigure}[b]{0.23\linewidth}
  \centering
  \includegraphics[width=0.8\linewidth]{images/intersense.jpg}
  Circular Data Matrix~\cite{NaimarkF02}
 \end{subfigure}\quad
 \begin{subfigure}[b]{0.14\linewidth}
 \centering
  \includegraphics[width=\linewidth]{images/pattKanji.pdf}
  ARToolkit~\cite{ARToolkit02}
 \end{subfigure}\quad
 \begin{subfigure}[b]{0.12\linewidth}
  \centering
  \includegraphics[width=\linewidth]{images/ARtag.jpg}
  ARTag\quad~\cite{Fiala05}
 \end{subfigure}\quad
 \begin{subfigure}[b]{0.12\linewidth}
  \centering
  \includegraphics[width=\linewidth]{images/pifiducial.jpg}
  PiTag\quad~\cite{Pitag13}
 \end{subfigure}\quad
 \begin{subfigure}[b]{0.14\linewidth}
  \centering
  \includegraphics[width=\linewidth]{images/our_fiducial}
  Our Fiducial
 \end{subfigure}
 \caption{Prior fiducial markers and our proposal.  While
 each code has its pros and cons depending on the environment, no
 prior code is expressly designed to be recognized under motion blur.}
 \label{fig:previous_work}
\end{figure}

Fiala~\cite{Fiala05} proposed a fiducial termed, ARTag, which is a
bi-tonal system consisting of a square border and an interior
6$\times$6 grid of black or white cells. The improvement of the ARTag
compared to the ARToolkit lies in the detection of corners instead of
detection of lines to find possible patterns.  This proved to be more
efficient than \cite{ARToolkit02} in terms of recognition rate as well
as the number of different patterns which can be created.  {\it The
reliance on both line and corner detection, however, hampers
recognition under motion blur.}

There were also attempts to use circular patterns instead of
rectangular.  Gatrell et al.~\cite{concentric} used concentric circles
for monocular pose estimation as well as object identification. Cho et
al.~\cite{Cho:2001,Cho97fastcolor} have used multicolor rings instead
of black and white rings~\cite{concentric} to increase possible number
of fiducials.  These multicolor rings are used in wide area tracking
in large scale applications.  {\it Although based on  concentric rings, these
approaches require the complete ring to be recognized; this is
impractical when the pattern undergoes directional motion blur.}

Naimark and Foxlin~\cite{NaimarkF02} proposed a circular bar code
called the Circular Data Matrix that is beneficial in terms of easy
detection and ability to have a large number of uniquely identifiable
codes.  To address the issue of occlusion, Bergamasco et
al.~\cite{runetag11} proposed the RUNE-tag fiducial by creating a
number of circular dots arranged in circular fashion. RUNE-tags can be
detected even when 50\% of the fiducial area is occluded. Bergamasco
et al.~\cite{Pitag13} proposed the PiTag fiducial, also composed of
circular structures but arranged in a rectangle, to exploit projective
invariant cross-ratio.  This provided similar occlusion resistance as
RUNE-tag but with even less circular dots. {\it All of these techniques,
however, rely on generic feature detection (e.g., circle detection)
and such algorithms break down under motion blur.}

@@

{\textbf{Tracking}}~~Fiducial detection between successive video
frames can be considered as a tracking problem where the tracked
object is the fiducial.  There is a very large body of research
dedicated to tracking and interested readers are referred
to~\cite{Yilmaz:2006} for a good survey.

Most tracking methods~\cite{Ross:2008,Wu:2009,Perez02,Mei:2009} assume
the image sequence to be blur free. In reality, however, the presence
of motion blur in a video sequence is often unavoidable. To this end,
Wu et al.~\cite{Wu:2011} proposed the Blur-driven Tracker (BLUT)
framework for tracking motion-blurred targets. BLUT is based on the
observation that although motion blur degrades the visual features of
the target, it may also provide useful cues about the movements to
help tracking.

The BLUT framework successfully tracks a blurred target when there is
uniform motion, and the position of the tracked object does not change
drastically in successive frames. However, as we will demonstrate in
Section~ 4, {\it the erratic motion from the quadcopters, as well as the
problem of dropped video frames makes the task too problematic for
BLUT to successfully track.}
%to be successively tracked.

@@

We begin by first motivating the need for a new blur resistant
fiducial by examining the performance of prior fiducials under motion
blur.  After this, we detail our design as well as the detection
algorithm used to find the fiducial in an image, or a video sequence.

\subsection{Prior Fiducials Under Motion Blur}\label{sec:blurtest}

Here we examine the the performance of two popular fiducials under
motion blur.  Specifically we examine ARTags ~\cite{Fiala05} and
PiTags~\cite{Pitag13} given their differences in geometric design and
the availability of an API to develop applications to recognize the tags.
\footnote{RUNE-tag~\cite{runetag11} currently does not provide access to  an
implementation to generate or recognize its tags.  The circular data
matrix~\cite{NaimarkF02} is available as a commercial product, however
it requires proprietary hardware.}

%To evaluate the ARTag and PiTag under blur, 
For controlled study, we simulate the appearance of the ARTag and PiTag
markers by scaling the tags to 150$\times$150 pixels.
Both fiducials are then blurred using linear motion blur at various
orientations with different  blur scales. The blur motion ranged from 15 to 50
in magnitude (measured in pixels), representing small to significant motion
blur. Figure~\ref{fig:artag_pitag} shows the visual appearance of the blurred
tags. We then try to detect the markers using the ALVAR library~\cite{alvar} and
PiTag library~\cite{ros_pitag}. The table in %the right side of
Figure~\ref{fig:artag_pitag} shows the recognition rate (in
percentage) of the two
fiducial markers at various blur scales over all  orientations.
As we can see, the PiTag performance quickly diminishes under small amounts of
blur, while at 35 units, the ARTag's recognition rate drops to less than
20\%.

\begin{figure}[t!]
\includegraphics[width=\linewidth]{images/artag_pitag.pdf}
%\captionof{figure}{Blurred ARTag and Pi--tag with various blur scales}
%\captionof{table}{Recognition rate of ARTag and Pi-tag fiducial marker atvarious blur scales}

\begin{tabularx}{\linewidth}{|Y|Y|Y|}
\cline{1-3}
\small{Blur} & \multicolumn{2}{c|}{ \small{Recognition Rate}}
\\\cline{2-3}
\small{Scale}& \small{PiTag} &	\small{ARTag} \\ \cline{1-3}
\small{15} & \small{100} & \small{100} \\ %\cline{1-3}
\small{30} & \small{0} & \small{100} \\  %\cline{1-3}
\small{35} & \small{0} & \small{19} \\ %\cline{1-3}
\small{50} & \small{0} & \small{0} \\ \cline{1-3}
\end{tabularx}
\captionof{figure}{\textbf{Top:} The PiTag and ARTag fiducials blurred
  with various blur scales at different orientations. \textbf{Bottom
    Table:} Recognition rate (in percent). We see that the recognition
  rates for both the tags are significantly reduced. For severe blur,
  detection fails.}
\label{fig:artag_pitag}
\end{figure}

@@

We propose a binary coded fiducial that uses concentric white rings of
equal widths on a black background with a blurred
border\footnote{Obviously, this design can be inverted to have a white
  background with black rings.}. The outermost and innermost rings
represent the start and end of the code and is embedded in the
fiducial.  The binary code is represented by the presence (or absence)
of rings between ``marker'' rings.

\begin{figure}
\centering
  \includegraphics[width=.24\linewidth]{images/newconcentric_00.pdf}
  \includegraphics[width=.24\linewidth]{images/newconcentric_01.pdf}
  \includegraphics[width=.24\linewidth]{images/newconcentric_10.pdf}
  \includegraphics[width=.24\linewidth]{images/newconcentric_11.pdf}
  \caption{Two bit, binary  fiducials, representing, from left to
    right 00, 01, 10, and 11.}
  \label{fig:fiducials}
\end{figure}

Depending on which ring is present or absent, the resulting binary
code will change. The number of different patterns depends on the
number of bits in the binary code. For example, if the binary code has
three bits, there will be a maximum of three rings between marker
rings and we end up with eight different patterns.
Figure~\ref{fig:fiducials} shows two bit binary fiducials.

Our fiducial detection strategy is different from
\cite{NaimarkF02,Pitag13} and works under significant amounts of blur.
As previously mentioned, our approach works under the observation that
the motion blur for the quadcopter's camera can be well modeled as
linear motion.  This linear motion blur assumption has been shown to
be reasonable in prior works~\cite{Moshe:2003,Moshe:2004}
targeting camera motion blur. Under this assumption, the scene
content perpendicular to the blur direction is unaffected by the blur.
Because of our circular design, the direction perpendicular to the
linear motion will still be recognizable as a linear pattern.
Figure~\ref{fig:blur_direction} shows examples of this using a pattern
under 
various motion directions.

\begin{figure}[h!]
\centering
\includegraphics[width=\linewidth]{images/blur_direction}
\caption{Change in blur direction changes the location and orientation
  of an unblurred linear pattern.}
\label{fig:blur_direction}
\end{figure}

\subsection{Detection Algorithm}

Figure~\ref{fig:overall_flow} shows the process involved in fiducial
detection. We give a brief overview of our algorithm here with each
step described in detail afterwards.  Our detection algorithm has four
steps. In Step~1, we apply a Gabor filter on the image to isolate the
potential locations of the pattern.  In Step~2, we find clusters of
patches in the Gabor output.  In Step~3, we perform Principal
Component Analysis (PCA) on each cluster to find the dominant
direction unaffected by blur.  Finally in Step~4, based on the
direction detected, we extract the intensity profile of the pattern
and classify the fiducial.

\begin{figure*}[ht!]
  \centering
  \includegraphics[width=0.8\linewidth]{images/overall_flow.pdf}
  \caption{An overview of our algorithm.
    The four step process includes (Step 1) Filtering,
    (Step 2) Component clustering, (Step 3) Dominant direction determination
    and (Step 4) Classification using prior training data.}
  \label{fig:overall_flow}
\end{figure*}

%\noindent Details of each of the steps are as follows:\\
\textbf{Gabor filter}~~A 2D Gabor filter is a Gaussian kernel function
modulated by a sinusoidal plane wave~\cite{Kruizinga:2002}. It is used to find
high gradient patches. In our case, it is used to detect portions
of the circular fiducials that were not affected by the blur.
We applied the Gabor filter in eight
different orientations ($\theta = 0, 45, 90, \ldots, 225, 270, 315$).  The
following parameters were used for creating each Gabor kernel: $\lambda$ (wavelength) $= 8$, $\gamma$
(aspect ratio) $= 0.5$, $\sigma$ (spread) $= 0.56\lambda$, $\psi$
(phase angle) $= 0$ (for real part), $\pi/2$ (for imaginary part).
Then $\ell^2$-norm of outputs along all orientations is calculated and finally
$\ell^2$-normalized image is thresholded with threshold set to $0.4$ (on scale of zero to
one).  For further details to the Gabor filter, see~\cite{Kruizinga:2002}.
%% MBS - is the Kruizinga filter resonable to cite?

\textbf{Clustering}~~The binarized Gabor filter responses are
treated as a set of connected components in the image.   Clustering
is used to find components that are located in a close spatial region.  We do
this via hierarchical clustering~\cite{ALGLIB} using unweighted
average linkage with a distance threshold set to 150.

\textbf{PCA}~~For each clustered set of components, we apply
PCA to determine the dominant direction.  This is done by examining
the orientation of the first principal component.  We then extract an
intensity profile patch in the input image along this direction as it
extends through the bounding box of the cluster. The signature of this
profile will be used to identify the code.

After finding the intensity profile, we project the pixel intensities,
and record the number and width of the white-to-black transitions.  If
the number of transitions, or the width of the transitions is not
consistent with what is allowable by our code, we reject the clustered
region as a potential fiducial.  Clustered components that have
allowable transitions counts and transitions with uniform widths are
further considered for classification to determine the binary code.

\textbf{Classification}~~As mentioned above, a small image patch
containing the intensity profile of the fiducial is used to identify
which of the many fiducials might be present and is represented by a
code.  We found that a training-based method using the k nearest
neighbor (k-NN) technique gave better results than trying to find
binary code directly by determining the presence (or absence) of ring
at particular positions in the image patch. This required
training-data which was easy to generate. A synthetic fiducial pattern
is blurred along 36 orientations (0, 10, 20, \ldots , 350) with blur
scale set to 40, and the intensity profile along the first principal
component from every output is taken as training data for that
fiducial pattern. Figure~\ref{fig:training_data} shows the process of
creating training data for fiducial with binary code ``01'' embedded
in it.  For a query image patch, we normalize the intensity range, and
then compare this information against the training data. The class
label from the closest top $K=5$ images in the training data is used
to label the patch.

\begin{figure}[h!]
\centering
  \includegraphics[width=0.95\linewidth]{images/training_data.pdf}
  \caption{Creating training data for the fiducial pattern with binary
  code ``01'': The synthetic pattern is blurred along various
  orientations and the intensity profile monitored and stored. The
  same process is used to create training data for all other fiducials.}
  \label{fig:training_data}
\end{figure}

@@

We have implemented our algorithm in C++ using the OpenCV library.
Experiments were performed on a laptop with Intel Core i7
processor(@3.4GHz) and 4GB RAM.  Source code to generate and recognize
the code, as well as the data sets used in this paper will be made
publicly available.
%Video clips of collected data are provided in the supplementary
%material.
% Meghshyam: Is it a desktop PC or a laptop?
